hierarchical data

### 2025-06-25 documentation/example.qmd
more documentation, clean up example

```{r}
fit1a <- fit_model_localhierarchy(runstep = "step1a",
                              hierarchical_level     =  hierarchical_level,
                              survey_df = dat,
                              chains = 4)

@@

```{r}
fit1a$post_summ <- get_posterior_summaries(fit1a)
fit1a$post_summ

@@

```{r}
fit_local <- fit_model_localhierarchy(runstep = "local_national",
                                  global_fit = fit1a,
                                  survey_df = dat,
                                  chains = 4)
```

@@

```{r}
res_global <- posterior_summary_hierparam(fit = fit1a, parname = "mu")
res_local <- posterior_summary_hierparam(fit = fit_local, parname = "mu")

@@

Plots that show how values have been fixed in the local runs

```{r}
bind_rows(res_local$subcluster %>% mutate(model = "local"),
          res_global$subcluster %>% mutate(model = "global")) %>%
  group_by(name, model) %>%
  reframe(y = val[quant == 0.5], ymin = val[quant == 0.025], ymax = val[quant == 0.975]) %>%
  ggplot(aes(y = y, x = name, color = model)) +
  geom_errorbar(aes(ymin = ymin, ymax = ymax)) +
  geom_point()

@@

And a plot to show the difference in estimates between global and local runs

```{r}
bind_rows(res_local$iso %>% mutate(model = "local"),
          res_global$iso %>% mutate(model = "global")) %>%
  group_by(name, model) %>%
  reframe(y = val[quant == 0.5], ymin = val[quant == 0.025], ymax = val[quant == 0.975]) %>%
  ggplot(aes(y = y, x = name, color = model)) +
  geom_errorbar(aes(ymin = ymin, ymax = ymax)) +
  geom_point()
```

@@

```{r}
hier_stan_data <- hierarchical_param_stan_data(
  global_fit = fit1a,
  param_name ="mu",
  param_data = hier_data,
  hierarchical_terms_fixed = hierarchical_level[1:2], 
  hierarchical_sigmas_fixed = hierarchical_level[1:3]
)
names(hier_stan_data)
```

```{r}
samples <- extract_parameter_subhierarchical(
        hierarchical_data = hier_data,
        subhierarchy = "iso",
        parname = "mu",
        fit_samples = fit1a$samples,
        morethan1param = FALSE
        )
samples
```

# Trace plots etc

here just checking the sigmas and first mu:

@@

fit <- fit1a
 parname_list <- list( c("mu_raw_estimate[1]",
                          "mu_sigma_estimate"))
#    plot_dirandname <- file.path(fit$output_dir, paste0(runname, "diagnostics.pdf"))
#    pdf(plot_dirandname, width = 11, height = 11)
    for (i in 1:length(parname_list)){
      parname <- parname_list[[i]]
      samp <- fit$samples$draws(parname )
      #mcmc_trace(samp)
      p <- mcmc_dens_overlay(samp)
      # for more than 1 parameter
      #mcmc_pairs(samp)
      print(p)
    }
#    dev.off()
```

>>>>

```{r}
fit_global <- fit_model_localhierarchy(runstep = "global_national",
                              hierarchical_level     =  hierarchical_level,
                              survey_df = dat,
                              chains = 4)

@@

```{r}
fit_global$post_summ <- get_posterior_summaries(fit_global)
fit_global$post_summ

@@

```{r}
fit_local <- fit_model_localhierarchy(runstep = "local_national",
                                  global_fit = fit_global,
                                  survey_df = dat,
                                  chains = 4)
```

@@

```{r}
res_global <- posterior_summary_hierparam(fit = fit_global, parname = "mu")
res_local <- posterior_summary_hierparam(fit = fit_local, parname = "mu")

@@

Plots that show outputs

```{r}
plot_posterior_summaries(res = res_global, modelname1 = "global")
```

how values have been fixed in the local runs

```{r}
plot_posterior_summaries(res = res_local, 
                         res2 = res_global,
                         modelname2 = "global", modelname1 = "local")

@@

selecting specific levels

```{r}
plot_posterior_summaries(res = res_local, 
                         res2 = res_global,
                         modelname2 = "global", modelname1 = "local",
                        hierarchy_select = "subcluster")

Level and selected areas

```{r}
plot_posterior_summaries(res = res_local, 
                         res2 = res_global,
                         modelname2 = "global", modelname1 = "local",
                        hierarchy_select = "iso",
                        areas_select = res_local$iso$name[1:100])

@@

```{r}
hier_stan_data <- hierarchical_param_stan_data(
  global_fit = fit_global,
  param_name = "mu",
  param_data = hier_data,
  hierarchical_terms_fixed = hierarchical_level[1:2], 
  hierarchical_sigmas_fixed = hierarchical_level[1:3]
)
names(hier_stan_data)
```

# Prior-posts

sigmas

```{r}
plot_prior_post_sigmas(fit = fit_globalsubnat, parname = "mu")

create plots for mu_raws

```{r}
plots <- plot_mu_raw(fit = fit_global, parname = "mu")
# a list with plots
```

summary plots gives summaries, 30 at a time

```{r}
plots[["summary_plots"]][[1]]

specific prior-post plots for one parameter are in plots\[\["plots_allmuraw"\]\]

```{r}
plots[["plots_allmuraw"]][[1]]
```

Simple trace plots etc

@@

fit <- fit_global
 parname_list <- list( c("mu_raw_estimate[1]",
                          "mu_sigma_estimate"))
#    plot_dirandname <- file.path(fit$output_dir, paste0(runname, "diagnostics.pdf"))
#    pdf(plot_dirandname, width = 11, height = 11)
    for (i in 1:length(parname_list)){
      parname <- parname_list[[i]]
      samp <- fit$samples$draws(parname )
      #mcmc_trace(samp)
      p <- mcmc_dens_overlay(samp)
      # for more than 1 parameter
      #mcmc_pairs(samp)
      print(p)
    }
#    dev.off()
```

### 2025-06-25 documentation/hierdata.qmd
more documentation, clean up example

```{r}
hier_stan_data$mu_n_terms # number of etas + global mu
length(unique(geo_unit_index$subcluster))
length(unique(geo_unit_index$iso))

@@

# Unsuccessful investigation into exposing stan build-in functions

>>>>

```{r}
hier_stan_data$mu_raw_n_terms # number of etas + global mu
length(unique(geo_unit_index$subcluster))
length(unique(geo_unit_index$iso))

@@

# Notes on unsuccessful investigation into exposing stan build-in functions
